### 2019-06-19 chapters/results/results.tex

\section{Overall Accuracy}
\begin{figure}
  \includegraphics[width=\linewidth]{images/rotation}
  \caption{MR: Rotation.}
  \label{fig:rotation 1}
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth]{images/shade}
  \caption{MR: Shading.}
  \label{fig: Shading}
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth]{images/sheer}
  \caption{MR: Sheering.}
  \label{fig: Sheer}
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth]{images/shift}
  \caption{MR: Shifting.}
  \label{fig: Shift}
\end{figure}

\section{Digit-by Accuracy}
\begin{figure}
  \includegraphics[width=\linewidth]{images/rotation1}
  \caption{MR: Rotation.}
  \label{fig:rotation 1}
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth]{images/shade1}
  \caption{MR: Shading.}
  \label{fig: Shading}
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth]{images/sheer1}
  \caption{MR: Sheering.}
  \label{fig: Sheer}
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth]{images/shift1}
  \caption{MR: Shifting.}
  \label{fig: Shift}
\end{figure}