### 2020-04-20 chapters/results/figures.tex

\subsection{Comparison of Accuracy by Transformation}\label{4.1.1}
In this section, we compare the accuracy of the original MNIST test data to the accuracy of the follow-up test cases. After generating the model and testing the original test data, we used the same model to test the follow-up transformed test dataset. The test datasets were applied to the models with the increasing value of transformations. 
As we discussed in section \ref{identifyingMR}, the follow-up test cases were generated by transforming the original dataset with an increasing value of metamorphic properties.  A copy of these new datasets was saved to serve as follow-up test cases. The follow-up test images have the same structure and labels as the original image. For example, the MNIST dataset was transformed by applying the shiftX metamorphic property for every pixel between $[-28,28]$. The models were then tested with transformed test data, and their accuracy was calculated for every follow-up test data. The following graphs show the change in accuracy of the whole dataset with the change in the degree of transformation. The X-axis shows the value of metamorphic transformation applied to the original test data, and the Y-axis shows the corresponding accuracy to that value of transformation.
    \begin{figure}[htb!]
    \centering
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{chapters/results/MT/Rotate.png}
        \caption{Rotation}
    \end{subfigure}
    \begin{subfigure}[b]{ 0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{chapters/results/MT/Shade.png}
        \caption{Shading}
    \end{subfigure}
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{chapters/results/MT/Shear.png}
        \caption{Shearing}
    \end{subfigure}
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{chapters/results/MT/ShiftX.png}
        \caption{ShiftX}
    \end{subfigure}
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
         \includegraphics[width=\textwidth]{chapters/results/MT/ShiftY.png}
        \caption{ShiftY}
    \end{subfigure}
    \caption{Accuracy of the algorithms.}
    \label{fig:Accuracy}
    \end{figure}

 From the graphs in \ref{fig:Accuracy}, we can see that different algorithm respond differently to the follow-up datasets. To quantify this difference, we calculated the area under the curve for all the graphs in \ref{fig:Accuracy}. The following table shows the area under each curve in \ref{fig:Accuracy}.\\
    \begin{table}[H]
    \centering
        \begin{tabular}{|c|c|c|c|c|c|}
        \hline
         & CNN & NN & NB & KNN & SVM \\
         \hline
        Rotate & 91.01 & 81.70 & 70.31 & 72.71 & 66.60  \\
        \hline
        Shade & 64.40 & 75.51 & 79.52 & 50.26 & 52.23 \\
        \hline
        Shear & 94.07 & 87.95 & 74.23 & 84.26 & 81.08 \\
        \hline
        ShiftX & 48.51 & 32.06 & 27.36 & 42.33 & 36.88 \\
        \hline
        ShiftY & 30.32 & 26.19 & 26.55 & 34.04 & 29.90 \\
        \hline
        \end{tabular}
        \caption{Area under curve.}
        \label{tbl:rot-file-format}
    \end{table}

    In table \ref{tbl:rot-file-format}, the convolutional neural network has the highest area under the curve for the rotate, shear, and shiftX property. This means that, on average, the accuracy of the convolutional neural network does not decrease as much as the other algorithms when tested with the various degree of transformed follow-up test cases.
    On the other hand, $k$NN has the lowest area under the curve for the shade metamorphic property. Thus, the accuracy of the $k$NN program drops rapidly as the value of shade property in the test dataset increases.
    % Figure 4.1(a) through (c) shows the variation in accuracy of . A higher area under curve means 
    % KNN has a higher accuracy but, the area under curve is much lower than the convolutional neural network and the 2-layer neural network. Suggests that if you want to use metamorphic transformation to reveal flaw in an algorithm some transformations may be preferable over others.

\subsection{Comparison of Accurracy by Class}\label{Digit-by Accuracy}
    % We discussed several papers in Section \ref{2.4MetamorphicTestingMachineLearning} on how the metamorphic relations affect datasets but none of them have investigated how the transformations affect the individual classes in those dataset. It is trivially easy to imagine that not all the data points in a dataset will be affected in the same way with all the transformations. For example, the digit $0$ will still be easily recognizable as $0$ even if it is rotated by 180\textdegree  due to it's symmetry. On the other hand, the digit $9$ can easily be confused with digit $6$ if rotated clockwise by 180\textdegree and vice-versa. Due to this non-uniform affect of transformations on the different data points we decided to investigate how each class of data points affects the accuracy of the model. Thus, besides studying the overall effect of the metamorphic transformations on the whole dataset, we will also study how they affect individual data classes in the MNIST dataset.
    We discussed several papers in Section \ref{2.4MetamorphicTestingMachineLearning} on the effect of metamorphic transformations on datasets, but no prior research was conducted on their impact on individual class in a dataset. It is trivially easy to imagine some classes not being affected in the same way in certain transformations. For example, the digit $0$ will still be easily recognizable as $0$, even if it is rotated by 180\textdegree  due to its symmetry. On the other hand, the digit $9$ has a higher chance of being mistaken for digit $6$ if rotated clockwise by 180\textdegree and vice-versa. Thus, in this section, we investigate the effect of metamorphic transformations on individual classes.

    % Figure \ref{fig:Digit by misclassification for Rotation MR} to \ref{fig:Digit by misclassification for ShiftY MR} shows the accuracy of MNIST dataset class labels $[0-9]$ for our metamorphic relations. Every figure in Figure 4.2-4.7 has ten sub-graphs, each, representing the correct classification of a class label. We want to see if there was a major difference in how each class is affected by a metamorphic relation to get a sense for robustness of the whole dataset. To quantify this we calculated the area under curve for each of the sub-graphs in tables \ref{tbl:test-file-formatRotate}-\ref{tbl:test-file-formatShiftY}. A higher area under curve means that the class was less influenced by the transformation with respect to other classes and it's accuracy on the transformed testing input will be closer to training data.
    Figure \ref{fig:Digit by misclassification for Rotation MR}-\ref{fig:Digit by misclassification for ShiftY MR} shows the accuracy of MNIST dataset class labels $[0-9]$ for our metamorphic relations. Every figure in \ref{fig:Digit by misclassification for Rotation MR}-\ref{fig:Digit by misclassification for ShiftY MR} has ten sub-graphs, each representing the correct classification of a class label with respect to the degree of transformation. We want to find clear outliers in the dataset which influence the accuracy of the model in a major way. To quantify this we calculated the area under curve for each of the sub-graphs in tables \ref{tbl:test-file-formatRotate}-\ref{tbl:test-file-formatShiftY}. A higher area under curve means that the transformation less influenced the class with respect to other classes, and its accuracy on the transformed testing input will be closer to training data.

    % Figure \ref{fig:Digit by misclassification for Rotation MR} shows the graphs for the accuracy of the classes in the MNIST dataset after applying the rotation transformation. There are five subplots in each sub-graph in Figure \ref{fig:Digit by misclassification for Rotation MR}. Each subplot represents the accuracy of that digit for the five machine learning algorithms we implemented. The X-axis represents the angle with which the test data points were rotated and the Y-axis represents the accuracy of all the test data points in that class.\\
    Figure \ref{fig:Digit by misclassification for Rotation MR} shows the graphs for the accuracy of the classes in the MNIST dataset after applying the rotation transformation. There are ten subgraphs, one for each class in $[0,9]$. Each sub-graph has five sub-plots, one for each machine learning algorithm that was implemented. The X-axis represents the degree of transformation (in this case, the angle of rotated), and the Y-axis represents the accuracy of all the test data points in that class.\\
    \begin{figure}[H]
    \centering
        \includegraphics[width=\textwidth]{chapters/results/MT/RotateAll.png}
        \caption{Class accuracy for MT:Rotate}
        \label{fig:Digit by misclassification for Rotation MR}
    \end{figure}

    % We did a similar analysis for all the metamorphic property identified in section \ref{identifyingMR}. For every class in the MNIST test dataset, we applied the transformations with increasing value and calculated the accuracy of the algorithms at each value. We also observed that all the algorithms started to predict the test data points as $1$ once the foreground became dark which is reflected in higher values of accuracy for digit $1$. \\
     We did a similar analysis for all the metamorphic property identified in section \ref{identifyingMR}. For every class in the MNIST test dataset, we applied the transformations with increasing value and calculated the accuracy of the algorithms at each value. We observe that all the test data is predicted as $1$ once the foreground becomes dark. This is also reflected in the higher values of accuracy in the graph for digit $1$.

    \begin{figure}[H]
    \centering
        \includegraphics[width=\textwidth]{chapters/results/MT/ShadeAll.png}
        \caption{Class accuracy for MT:Shading}
        \label{fig:Digit by misclassification for Shading MR}
    \end{figure}

    The following figures show similar graphs for the metamorphic properties shear, shifting the images on X axis, shifting the images on Y axis.

    \begin{figure}[H]
    \centering
        \includegraphics[width=\textwidth]{chapters/results/MT/ShearAll.png}
        \caption{Class accuracy for MT:Shear}
        \label{fig:Digit by misclassification for Shear MR}
    \end{figure}

    \begin{figure}[ht!]
    \centering
        \includegraphics[width=\textwidth]{chapters/results/MT/ShiftXAll.png}
        \caption{Class accuracy for MT: ShiftX}
        \label{fig:Digit by misclassification for ShiftX MR}
    \end{figure}

    \begin{figure}[H]
    \centering
        \includegraphics[width=\textwidth]{chapters/results/MT/ShiftYAll.png}
        \caption{Class accuracy for MT: ShiftY}
        \label{fig:Digit by misclassification for ShiftY MR}
    \end{figure}

  To further compare the accuracy of the different classes, we calculated the area under each curve for each plot. The following tables show the area under the curve for each class $[0,9]$ in the above graphs. The cells with \textcolor{green}{green} background has the highest value among all the cells in that row, and the ones in \textcolor{red}{red} has the lowest.
    \begin{table}[H]
    \centering
        \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|}
        \hline
         & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 \\
        \hline
        CNN & \cellcolor{green!25} 97.01 & 87.69 & 92.87 & 90.29 & 87.77 & 91.86 & 96.54 & 90.47 & 88.99 & \cellcolor{red!25}86.62 \\ 
        \hline
        NN &   94.34 & 84.75 & 75.98 & 79.87 & 80.47 & \cellcolor{red!25}68.33 & \cellcolor{green!25}95.66 & 78.64 & 76.16 & 82.76 \\
        \hline
        NB & \cellcolor{green!25}88.80 & 76.64 & 64.08 & 75.23 & 66.16 & \cellcolor{red!25}56.60 & 84.09 & 57.65 & 65.20 & 68.69 \\
        \hline
        KNN & \cellcolor{green!25}98.31 & 90.80 & 65.39 & 81.12 & \cellcolor{red!25}56.70 & 73.24 & 80.56 & 60.03 & 62.05 & 58.92 \\
        \hline
        SVM & \cellcolor{green!25}93.97 & 76.54 & \cellcolor{red!25}50.38 & 60.45 & 76.49 & 51.74 & 79.87 & 61.49 & 61.58 & 53.49 \\
        \hline
        \end{tabular}
        \caption{Area under curve for each class for MR: Rotation.}
        \label{tbl:test-file-formatRotate}
    \end{table}

     In table \ref{tbl:test-file-formatRotate} we can see that the digit $0$ has the highest area under the curve in the convolutional neural network implementation and the digit $9$ has the lowest area. From this, we can say that the accuracy of the sub-set of the MNIST dataset with class label $0$ is not affected as much as the other classes when the rotation property is applied. Thus, when creating follow-up test cases for a convolutional neural network, the tester can apply the rotation property to the digit $0$ to a larger angle than any other classes. We will make these recommendations more concrete in \ref{tbl:indexrecommendations}.
     We generated the following tables to perform the same analysis on all the metamorphic properties.

    \begin{table}[H]
    \centering
    \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|}
    \hline
     & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 \\
    \hline
    CNN & 63.02 & \cellcolor{green!25}99.38 & 59.26 & 60.58 & 69.07 & 58.96 & 60.93 & 63.65 & \cellcolor{red!25}53.64 & 55.51 \\ 
    \hline
    NN & 72.64 & \cellcolor{green!25}96.66 & 72.61 & 69.35 & 80.50 & 80.77 & 81.16 & 82.36 & \cellcolor{red!25}50.98 & 68.08 \\
    \hline
    NB & 89.85 & \cellcolor{green!25}93.71 & 75.36 & 85.34 & 73.75 & \cellcolor{red!25}67.04 & 88.28 & 71.61 & 70.71 & 79.58 \\
    \hline
    KNN & 52.81 & \cellcolor{green!25}99.87 & 39.91 & 46.68 & 42.49 & 48.36 & 47.00 & 52.17 & \cellcolor{red!25}32.81 & 40.52 \\
    \hline
    SVM & 47.12 & \cellcolor{green!25}99.61 & 47.91 & 44.62 & 53.40 & 56.73 & 47.62 & 52.26 & \cellcolor{red!25}31.80 & 41.23 \\
    \hline
    \end{tabular}
    \caption{Area under curve for each class for MR: Shading.}
    \label{tbl:test-file-formatShade}
    \end{table}
    In table \ref{tbl:test-file-formatShade}, we see that the digit $1$ had the highest accuracy as compared to every other class in the dataset for every algorithm we implemented. On the other hand, the digit $8$ had the lowest accuracy in four out of the five algorithms we implemented.
    \begin{table}[H]
    \centering
    \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|}
    \hline
    & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 \\
    \hline
    CNN & 94.75 & 95.80 & 96.42 & 91.66 & 94.39 & 96.26 &\cellcolor{green!25}98.05 & 94.18 & \cellcolor{red!25}88.89 & 90.28  \\ 
    \hline
    NN & 90.99 & 90.21 & 88.38 & 87.06 & 85.43 & \cellcolor{red!25}82.57 & \cellcolor{green!25}95.77 & 88.55 & 86.13 & 84.41 \\
    \hline
    NB & 79.97 & 79.09 & 78.67 & \cellcolor{green!25}81.60 & 70.81 & 68.39 & 79.37 & 72.12 & 69.02 & \cellcolor{red!25}63.31 \\
    \hline
    KNN & 96.32 & \cellcolor{green!25}99.54 & 85.20 & 88.43 & 78.91 & 83.46 & 93.39 & \cellcolor{red!25}69.79 & 76.79 & 70.73 \\
    \hline
    SVM & \cellcolor{green!25}90.17 & 90.03 & 79.92 & 78.93 & 85.78 & 76.11 & 88.71 & 75.08 & 82.16 & \cellcolor{red!25}63.90 \\
    \hline
    \end{tabular}
    \caption{Area under curve for each class for MR: Shear.}
    \label{tbl:test-file-formatShear}
    \end{table}
    In table \ref{tbl:test-file-formatShear}, we find that the digit $6$ had the highest area under curve while the digits $8$ and $5$ have the lowest area under curve for the convolutional neural network and the seven-layer neural network. The digits $3$, $1$, and $0$ have the highest area under curve in the Naive Bayes, KNN, and, SVM implementations respectively.
    \begin{table}[htb!]
    \centering
    \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|}
    \hline
    & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 \\
    \hline
    CNN & \cellcolor{red!25}25.17 & \cellcolor{green!25}92.04 & 65.01 & 31.49 & 64.78 & 42.84 & 46.12 & 61.14 & 27.62 & 28.84 \\
    \hline
    NN & 21.30 & \cellcolor{red!25}15.45 & 48.59 & 30.03 & 31.99 & 33.89 & 40.76 & \cellcolor{green!25}49.62 & 21.58 & 27.42 \\
    \hline
    NB & 19.03 & 14.50 & \cellcolor{green!25}60.03 & 27.11 & 30.67 & 45.91 & 20.14 & 25.42 & 16.63 & \cellcolor{red!25}14.14 \\
    \hline
    KNN & 27.56 & \cellcolor{green!25}67.33 & 56.41 & 39.62 & 32.09 & 58.53 & 34.05 & 50.71 & 32.65 & \cellcolor{red!25}24.31 \\
    \hline
    SVM & \cellcolor{red!25}22.05 & 40.13 & \cellcolor{green!25}57.64 & 33.38 & 39.98 & 49.06 & 35.29 & 43.62 & 24.94 & 22.74 \\
    \hline
    \end{tabular}
    \caption{Area under curve for each class for MR: ShiftX.}
    \label{tbl:test-file-formatShiftX}
    \end{table}

    \begin{table}[H]
    \centering
    \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|}
    \hline
    & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 \\
    \hline
    CNN & 26.80 & \cellcolor{green!25}52.91 & 25.32 & 22.45 & 30.47 & 26.39 & 38.99 & 29.25 & \cellcolor{red!25}21.09 & 29.53 \\ 
    \hline
    NN &  20.99 & 35.29 & 22.58 & 14.69 & 20.49 & \cellcolor{red!25}14.18 & \cellcolor{green!25}52.86 & 30.59 & 15.52 & 34.77 \\
    \hline
    NB & 19.95 & \cellcolor{green!25}56.37 & 23.13 & 19.51 & 20.90 & \cellcolor{red!25}16.07 & 36.29 & 23.58 & 24.89 & 24.81 \\
    \hline
    KNN & 28.72 & \cellcolor{green!25}95.26 & 26.83 & 23.77 & 36.41 & 25.06 & 34.64 & 29.50 & \cellcolor{red!25}19.31 & 20.93 \\
    \hline
    SVM & 25.09 & \cellcolor{green!25}61.47 & 26.25 & \cellcolor{red!25}17.09 & 30.53 & 25.75 & 43.48 & 31.59 & 17.84 & 19.90  \\
    \hline
    \end{tabular}
    \caption{Area under curve for each class for MR: ShiftY.}
    \label{tbl:test-file-formatShiftY}
    \end{table}
    For the metamorphic transformation shiftY, we find that the digit $1$ has the highest area under the curve for all but the seven-layer neural network implementation. Thus, images with class $1$ can be shifted to a larger value before they get confused with other classes.

A similar analysis was also done for the Fashion-MNIST, and, EMNIST dataset but due to space limitation they are not reported here, instead, they are reported on the project's github page \cite{abhishek2020}.

\section{Recommendations}
% In the tables 4.2 to 4.6 we have seen that the different classes are affected differently depending upon the algorithm and the metamorphic transformation applied to them. In this section, we want to make recommendations to pick the best values for the metamorphic properties such that the the accuracy of the transformed dataset will be close to the original dataset. For our purpose we define accuracy being close as it being within $90\%$ of the origin dataset. To do this we first calculated the accuracy of each class in the original dataset without applying any transformation. We then calculated the degrees of transformation for which the accuracy of the transformed dataset was greater than or equal to $90\%$ of the accuracy of the original dataset for the five algorithms. Based on this exercise we make the recommendation to testers who want to use these properties to generate follow-up test cases. The following table shows the minimum and maximum values of transformations: 
In tables \ref{tbl:test-file-formatRotate} to \ref{tbl:test-file-formatShiftY}, we have seen that the different classes are affected differently depending upon the algorithm, and the metamorphic transformation applied to them. In this section, we want to make recommendations to pick the best values for transforming metamorphic properties such that the accuracy of the transformed dataset will be close to the original dataset. For this framework, we define accuracy being close as it is within $90\%$ of the original dataset. A tester will be able to use these recommendations to generate new data that are similar to the original dataset.
To make the recommendations, we first calculated the accuracy of each class in the original dataset without applying any transformation. We then calculated the degrees of transformation for which the transformed dataset was similar to the original dataset by testing them on five algorithms. Since different algorithms had different values for the range of transformations, we took the minimum range from all the algorithms.
Based on this exercise, we make the recommendation to testers who want to use these properties to generate follow-up test cases. The table \ref{tbl:indexrecommendations} shows the minimum and maximum values of transformations for each class of data in the MNIST dataset which will produce a dataset similar to the origin dataset: 
% where the accuracy was found to be within $90\%$ of the accuracy without any transformation.

% \pgfplotstableread[row sep=\\,col sep=&]{
%   Class & RotateX & RotateY & ShadeX & ShadeY & ShearX & ShearY & ShiftXx & ShiftXy    & ShiftYx & ShiftYy\\
%     0 &     0 &     35 &         0.1 &     51      & 0.2 & 22    & 0.3        & 3         & 0.4       & 2  \\
%     1 &     200 &     22 &        200.1 &     99      & 200.2 & 20    & 200.3        & -25       & 200.4       & 3  \\
%     2 &     400 &     12 &        400.1 &     50      & 400.2 & 26    & 400.3        & -24       & 400.4      & 2  \\
%     3 &     600 &     13 &        600.1 &     51      & 600.2 & 22    & 600.3        & 1         & 600.4       & 2  \\
%     4 &     800 &     15 &        800.1 &     51      & 800.2 & 19    & 800.3        & 4         & 800.4       & 1  \\
%     5 &     1000 &     10 &        1000.1 &     50      & 1000.2 & 11    & 1000.3        & 6         & 1000.4       & 2  \\
%     6 &     1200 &     37 &        1200.1 &     50      & 1200.2 & 30    & 1200.3        & 3         & 1200.4       & -27  \\
%     7 &     1400 &     11 &        1400.1 &     51      & 1400.2 & 16    & 1400.3        & 2         & 1400.4        & 2  \\
%     8 &     1600 &     15 &        1600.1 &     50      & 1600.2 & 21    & 1600.3        & 3         & 1600.4       & 2  \\
%     9 &     1800 &     16 &        1800.1 &     50      & 1800.2 & 15    & 1800.3        & 2         & 1800.4       & 2  \\
% }\recomx

% \pgfplotstableread[row sep=\\,col sep=&]{
%   Class & RotateY & RotateX & ShadeY & ShadeX & ShearY & ShearX & ShiftXy & ShiftXx    & ShiftYy & ShiftYx\\
%     0   & -45       & 0     & 0         & 0.1    & -35  & 0.2    & -3       & 0.3       & -2    &  0.4  \\
%     1   & -19       & 200     & 0         & 200.1    & -28  & 200.2    & 16       & 200.3         & -5  & 200.4  \\
%     2   & -14       & 400    & 0        & 400.1   & -25  & 400.2    & -4      & 400.3       & -2    & 400.4  \\
%     3   & -17       & 600    & 0         & 600.1    & -25  & 600.2    & -4     & 600.3         & -2  & 600.4  \\
%     4   & -14       & 800     & 0         & 800.1    & -21  & 800.2    & -3       & 800.3         & -3  & 800.4  \\
%     5   & -11       & 1000     & 0         & 1000.1    & -25  & 1000.2    & -1       & 1000.3         & -1  & 1000.4  \\
%     6   & -16       & 1200     & 0         & 1200.1    & -18  & 1200.2    & -3       & 1200.3         & -3  & 1200.4  \\
%     7   & -19       & 1400     & 0         & 1400.1    & -24  & 1400.2    & -4       & 1400.3         & -1  & 1400.4  \\
%     8   & -14       & 1600     & 0         & 1600 .1    & -32  & 1600 .2    & -1       & 1600 .3         & -2  & 1600 .4  \\
%     9   & -11       & 1800     & 0         & 1800.1    & -12  & 1800.2    & -2       & 1800.3         & -1  & 1800.4 \\
% }\recomy

% \begin{tikzpicture}

% \begin{groupplot}[group style={group size= 1 by 2},xtick=\empty]%[ybar stacked,xtick=\empty,]%ytick=\empty]
% \nextgroupplot[ybar, bar width=.1cm,xtick=\empty,yshift=1cm, ymin=0,ymax=60,legend style = {at={(1.05, 0.5)}, anchor = west, legend columns =1, draw=none, area legend},  ymajorgrids = true,width=\textwidth,height=.5\textwidth,]%ytick=\empty]
% \addplot[fill=blue,draw=black] table[x=RotateX,y=RotateY]{\recomx};
% \addplot[fill=red,draw=black] table[x=ShadeX,y=ShadeY]{\recomx};
% \addplot[fill=green,draw=black] table[x=ShearX,y=ShearY]{\recomx};
% \addplot[fill=yellow,draw=black] table[x=ShiftXx,y=ShiftXy]{\recomx};
% \addplot[fill=brown,draw=black] table[x=ShiftYx,y=ShiftYy]{\recomx};

% \addlegendentry{CNN}
% \addlegendentry{NN}
% \addlegendentry{NB}
% \addlegendentry{SVM}
% \addlegendentry{KNN}

% \nextgroupplot[ybar, width=\textwidth,height=.5\textwidth, bar width=.1cm,yshift=1cm,ymin=-60,ymax=0,legend style = {at={(1.05, 1.65)}, anchor = west, legend columns =1, draw=none, area    legend},ymajorgrids = true,xtick=data, xticklabels from table={\recomy}{Class}, xticklabel style={yshift=-2ex, xshift=-2ex,anchor= west},]
% \addplot[fill=blue,draw=black] table[x=RotateX,y=RotateY]{\recomy};
% \addplot[fill=red,draw=black] table[x=ShadeX,y=ShadeY]{\recomy};
% \addplot[fill=green,draw=black] table[x=ShearX,y=ShearY]{\recomy};
% \addplot[fill=yellow,draw=black] table[x=ShiftXx,y=ShiftXy]{\recomy};
% \addplot[fill=brown,draw=black] table[x=ShiftYx,y=ShiftYy]{\recomy};

% \end{groupplot}
% \end{tikzpicture}

\begin{center}
 \begin{table}[H]
  \centering
    \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}
    \hline
     \multirow{2}{*}{Class} & \multicolumn{2}{c|}{Rotation} & \multicolumn{2}{c|}{Shade} & \multicolumn{2}{c|}{Shear} & \multicolumn{2}{c|}{ShiftX} & \multicolumn{2}{c|}{ShiftY} \\\cline{2-11} 
     & Min & Max & Min & Max & Min & Max & Min & Max & Min & Max \\
     \hline
    $0$ & $-45$ & $35$ & $0$ & $51$ & $-35$ & $22$ & $-3$ & $3$ & $-2$ & $2$  \\
    $1$ & $-19$ & $22$ & $0$ & $99$ & $-28$ & $20$ & $-2$ & $2$ & $-5$ & $3$  \\
    $2$ & $-14$ & $12$ & $0$ & $50$ & $-25$ & $26$ & $-4$ & $7$ & $-2$ & $2$  \\
    $3$ & $-17$ & $13$ & $0$ & $51$ & $-25$ & $22$ & $-4$ & $1$ & $-2$ & $2$  \\
    $4$ & $-14$ & $15$ & $0$ & $51$ & $-21$ & $19$ & $-3$ & $4$ & $-3$ & $1$  \\
    $5$ & $-11$ & $10$ & $0$ & $50$ & $-25$ & $11$ & $-1$ & $6$ & $-1$ & $2$  \\
    $6$ & $-16$ & $37$ & $0$ & $50$ & $-18$ & $30$ & $-3$ & $3$ & $-3$ & $2$  \\
    $7$ & $-19$ & $11$ & $0$ & $51$ & $-24$ & $16$ & $-4$ & $2$ & $-1$ & $2$  \\
    $8$ & $-14$ & $15$ & $0$ & $50$ & $-32$ & $21$ & $-1$ & $3$ & $-2$ & $2$  \\
    $9$ & $-11$ & $16$ & $0$ & $50$ & $-12$ & $15$ & $-2$ & $2$ & $-1$ & $2$  \\
    \hline
    \end{tabular}
    \caption{Recommendations for using the transformed MNIST dataset.}
    \label{tbl:indexrecommendations}
   \end{table}   
\end{center}
We have also provided these recommendations for the Fashion-MNIST and EMNIST datasets on the project's github page \cite{abhishek2020}.
% \subsection{Misclassification of digits}
%     In order to see how each digit is affected by changes in  mp. For each digit 0-9 we calculated the number of correct predictions made by the model as compared to total number of .