### 2020-04-22 chapters/results/robustnessGraphs.tex

\pgfplotstableread[row sep=\\,col sep=&]{
    Mx & My & Fx & Fy & Ex & Ey \\
    0 & 39 & 0.1 & 13 & 0.2 & 24 \\
    4 & 27 & 4.1 & 19 & 4.2 & 19 \\
    8 & 25 & 8.1 & 19 & 8.2 & 20 \\
    12 & 16 & 12.1 & 18 & 12.2 & 21 \\
    16 & 19 & 16.1 & 15 & 16.2 & 24 \\
    }\mnistx

\pgfplotstableread[row sep=\\,col sep=&]{
   Algo & Mx & My & Fx & Fy & Ex & Ey \\
   CNN & 0 & -34 & 0.1 & -14 & 0.2 & -20 \\
    NN & 4 & -28 & 4.1 & -17 & 4.2 & -18 \\
    NB & 8 & -24 & 8.1 & -19 & 8.2 & -17 \\
    SVM & 12 & -16 & 12.1 & -17 & 12.2 & -20 \\
     KNN & 16 & -18 & 16.1 & -16 & 16.2 & -25 \\
    }\mnisty

\begin{figure}[H]

\begin{tikzpicture}

\begin{groupplot}[group style={group size= 1 by 2},xtick=\empty]%[ybar stacked,xtick=\empty,]%ytick=\empty]
\nextgroupplot[ylabel={Range of transformations},ybar, bar width=.2cm,xtick=\empty,yshift=1cm, ymin=0,ymax=100,legend style = {at={(1.05, 0.5)}, anchor = west, legend columns =1, draw=none, area legend},  ymajorgrids = true,width=0.85\textwidth,height=.35\textwidth, y label style={at={(axis description cs:-0.01,0.001)}},]%ytick=\empty]
\addplot[fill=blue,draw=black] table[x=Mx,y=My]{\mnistx};
\addlegendentry{MNIST}

\addplot[fill=red,draw=black] table[x=Fx,y=Fy]{\mnistx};
\addlegendentry{Fashion}

\addplot[fill=brown,draw=black] table[x=Ex,y=Ey]{\mnistx};
\addlegendentry{EMNIST}

\nextgroupplot[ybar, width=0.85\textwidth,height=.35\textwidth, bar width=.2cm,yshift=1cm,ymin=-100,ymax=0,legend style = {at={(1.05, 1.65)}, anchor = west, legend columns =1, draw=none, area    legend},ymajorgrids = true,xtick=data, xticklabels from table={\mnisty}{Algo}, xticklabel style={yshift=-2ex, xshift=-2ex,anchor= west},xlabel={Algorithms},x label style={at={(axis description cs:0.5,-0.1)},anchor=north}]
\addplot[fill=blue,draw=black] table[x=Mx,y=My]{\mnisty};
\addplot[fill=red,draw=black] table[x=Fx,y=Fy]{\mnisty};
\addplot[fill=brown,draw=black] table[x=Ex,y=Ey]{\mnisty};
\end{groupplot}
\end{tikzpicture}
\caption{Robustness graph for rotation property}
\end{figure}

\pgfplotstableread[row sep=\\,col sep=&]{
    Mx & My & Fx & Fy & Ex & Ey \\
    0 & 55 & 0.1 & 55 & 0.2 & 53 \\
    4 & 69 & 4.1 & 53 & 4.2 & 61 \\
    8 & 96 & 8.1 & 76 & 8.2 & 92 \\
    12 & 52 & 12.1 & 52 & 12.2 & 51 \\
    16 & 51 & 16.1 & 52 & 16.2 & 51 \\
    }\shadex

\pgfplotstableread[row sep=\\,col sep=&]{
   Algo & Mx & My & Fx & Fy & Ex & Ey \\
    CNN & 0 & 0 & 0.1 & 0 & 0.2 & 0 \\
    NN & 4 & 0 & 4.1 & 0 & 4.2 & 0 \\
    NB & 8 & 0 & 8.1 & 0 & 8.2 & 0 \\
    SVM & 12 & 0 & 12.1 & 0 & 12.2 & 0 \\
    KNN & 16 & 0 & 16.1 & 0 & 16.2 & 0 \\
    }\shadey

\begin{figure}[H]
\begin{tikzpicture}
\begin{groupplot}[group style={group size= 1 by 2},xtick=\empty]%[ybar stacked,xtick=\empty,]%ytick=\empty]
\nextgroupplot[ylabel={Range of transformations},ybar, bar width=.2cm,xtick=\empty,yshift=1cm, ymin=0,ymax=100,legend style = {at={(1.05, 0.5)}, anchor = west, legend columns =1, draw=none, area legend},  ymajorgrids = true,width=0.85\textwidth,height=.5\textwidth, y label style={at={(axis description cs:-.009,.5)}},]%ytick=\empty]
\addplot[fill=blue,draw=black] table[x=Mx,y=My]{\shadex};
\addlegendentry{MNIST}

\addplot[fill=red,draw=black] table[x=Fx,y=Fy]{\shadex};
\addlegendentry{Fashion}

\addplot[fill=brown,draw=black] table[x=Ex,y=Ey]{\shadex};
\addlegendentry{EMNIST}

\nextgroupplot[ybar, width=0.85\textwidth,height=0, bar width=.2cm,yshift=1cm,ymin=-100,ymax=0,legend style = {at={(1.05, 1.65)}, anchor = west, legend columns =1, draw=none, area    legend},ymajorgrids = true,xtick=data, xticklabels from table={\shadey}{Algo}, xticklabel style={yshift=-2ex, xshift=-2ex,anchor= west},xlabel={Algorithms},x label style={at={(axis description cs:0.5,-0.1)},anchor=north}]
\addplot[fill=blue,draw=black] table[x=Mx,y=My]{\shadey};
\addplot[fill=red,draw=black] table[x=Fx,y=Fy]{\shadey};
\addplot[fill=brown,draw=black] table[x=Ex,y=Ey]{\shadey};

\end{groupplot}
\end{tikzpicture}
\caption{Robustness graph for shading property}
\end{figure}

\pgfplotstableread[row sep=\\,col sep=&]{
    Mx & My & Fx & Fy & Ex & Ey \\
    0 & 47 & 0.1 & 24 & 0.2 & 30 \\
    4 & 43 & 4.1 & 29 & 4.2 & 24 \\
    8 & 34 & 8.1 & 33 & 8.2 & 26 \\
    12 & 23 & 12.1 & 31 & 12.2 & 29 \\
    16 & 23 & 16.1 & 29 & 16.2 & 35 \\
    }\shearx

\pgfplotstableread[row sep=\\,col sep=&]{
   Algo & Mx & My & Fx & Fy & Ex & Ey \\
    CNN & 0 & -48 & 0.1 & -25 & 0.2 & -30 \\
    NN & 4 & -46 & 4.1 & -30 & 4.2 & -26 \\
    NB & 8 & -31 & 8.1 & -43 & 8.2 & -23 \\
    SVM & 12 & -28 & 12.1 & -40 & 12.2 & -33 \\
    KNN & 16 & -31 & 16.1 & -32 & 16.2 & -45 \\
    }\sheary

\begin{figure}[H]
\begin{tikzpicture}
\begin{groupplot}[group style={group size= 1 by 2},xtick=\empty]%[ybar stacked,xtick=\empty,]%ytick=\empty]
\nextgroupplot[ylabel={Range of transformations},ybar, bar width=.2cm,xtick=\empty,yshift=1cm, ymin=0,ymax=100,legend style = {at={(1.05, 0.5)}, anchor = west, legend columns =1, draw=none, area legend},  ymajorgrids = true,width=0.85\textwidth,height=.35\textwidth, y label style={at={(axis description cs:-0.01,0.001)}},]%ytick=\empty]
\addplot[fill=blue,draw=black] table[x=Mx,y=My]{\shearx};
\addlegendentry{MNIST}

\addplot[fill=red,draw=black] table[x=Fx,y=Fy]{\shearx};
\addlegendentry{Fashion}

\addplot[fill=brown,draw=black] table[x=Ex,y=Ey]{\shearx};
\addlegendentry{EMNIST}

\nextgroupplot[ybar, width=0.85\textwidth,height=.35\textwidth, bar width=.2cm,yshift=1cm,ymin=-100,ymax=0,legend style = {at={(1.05, 1.65)}, anchor = west, legend columns =1, draw=none, area    legend},ymajorgrids = true,xtick=data, xticklabels from table={\sheary}{Algo}, xticklabel style={yshift=-2ex, xshift=-2ex,anchor= west},xlabel={Algorithms},x label style={at={(axis description cs:0.5,-0.1)},anchor=north}]
\addplot[fill=blue,draw=black] table[x=Mx,y=My]{\sheary};
\addplot[fill=red,draw=black] table[x=Fx,y=Fy]{\sheary};
\addplot[fill=brown,draw=black] table[x=Ex,y=Ey]{\sheary};

\end{groupplot}
\end{tikzpicture}
\caption{Robustness graph for Shear property}
\end{figure}

\pgfplotstableread[row sep=\\,col sep=&]{
    Mx & My & Fx & Fy & Ex & Ey \\
    0 & 5 & 0.1 & 1 & 0.2 & 1 \\
    4 & 3 & 4.1 & 1 & 4.2 & 1 \\
    8 & 3 & 8.1 & 2 & 8.2 & 1 \\
    12 & 4 & 12.1 & 1 & 12.2 & 1 \\
    16 & 4 & 16.1 & 1 & 16.2 & 1 \\
    }\shearxX

\pgfplotstableread[row sep=\\,col sep=&]{
   Algo & Mx & My & Fx & Fy & Ex & Ey \\
    CNN & 0 & -5 & 0.1 & -1 & 0.2 & -1 \\
    NN & 4 & -3 & 4.1 & -1 & 4.2 & -1 \\
    NB & 8 & -3 & 8.1 & -2 & 8.2 & -1 \\
    SVM & 12 & -4 & 12.1 & -1 & 12.2 & -1 \\
    KNN & 16 & -5 & 16.1 & -1 & 16.2 & -1 \\
    }\shearxY

\begin{figure}[H]
\begin{tikzpicture}
\begin{groupplot}[group style={group size= 1 by 2},xtick=\empty]%[ybar stacked,xtick=\empty,]%ytick=\empty]
\nextgroupplot[ylabel={Range of transformations},ybar, bar width=.2cm,xtick=\empty,yshift=1cm, ymin=0,ymax=10,legend style = {at={(1.05, 0.5)}, anchor = west, legend columns =1, draw=none, area legend},  ymajorgrids = true,width=0.85\textwidth,height=.35\textwidth, y label style={at={(axis description cs:-0.01,0.001)}},]%ytick=\empty]
\addplot[fill=blue,draw=black] table[x=Mx,y=My]{\shearxX};
\addlegendentry{MNIST}

\addplot[fill=red,draw=black] table[x=Fx,y=Fy]{\shearxX};
\addlegendentry{Fashion}

\addplot[fill=brown,draw=black] table[x=Ex,y=Ey]{\shearxX};
\addlegendentry{EMNIST}

\nextgroupplot[ybar, width=0.85\textwidth,height=.35\textwidth, bar width=.2cm,yshift=1cm,ymin=-10,ymax=0,legend style = {at={(1.05, 1.65)}, anchor = west, legend columns =1, draw=none, area    legend},ymajorgrids = true,xtick=data, xticklabels from table={\shearxY}{Algo}, xticklabel style={yshift=-2ex, xshift=-2ex,anchor= west},xlabel={Algorithms},x label style={at={(axis description cs:0.5,-0.1)},anchor=north}]
\addplot[fill=blue,draw=black] table[x=Mx,y=My]{\shearxY};
\addplot[fill=red,draw=black] table[x=Fx,y=Fy]{\shearxY};
\addplot[fill=brown,draw=black] table[x=Ex,y=Ey]{\shearxY};
\end{groupplot}
\end{tikzpicture}
\caption{Robustness graph for ShiftX property}
\end{figure}

\pgfplotstableread[row sep=\\,col sep=&]{
    Mx & My & Fx & Fy & Ex & Ey \\
    0 & 3 & 0.1 & 1 & 0.2 & 1 \\
    4 & 2 & 4.1 & 1 & 4.2 & 0 \\
    8 & 2 & 8.1 & 1 & 8.2 & 1 \\
    12 & 3 & 12.1 & 1 & 12.2 & 1 \\
    16 & 3 & 16.1 & 1 & 16.2 & 1 \\
    }\shearyX

\pgfplotstableread[row sep=\\,col sep=&]{
   Algo & Mx & My & Fx & Fy & Ex & Ey \\
    CNN & 0 & -4 & 0.1 & -1 & 0.2 & -1 \\
    NN & 4 & -2 & 4.1 & 0 & 4.2 & -1 \\
    NB & 8 & -3 & 8.1 & -1 & 8.2 & -1 \\
    SVM & 12 & -3 & 12.1 & -1 & 12.2 & -1 \\
    KNN & 16 & -4 & 16.1 & -1 & 16.2 & -1 \\
    }\shearyY

\begin{figure}[H]
\begin{tikzpicture}
\begin{groupplot}[group style={group size= 1 by 2},xtick=\empty]%[ybar stacked,xtick=\empty,]%ytick=\empty]
\nextgroupplot[ylabel={Range of transformations},ybar, bar width=.2cm,xtick=\empty,yshift=1cm, ymin=0,ymax=10,legend style = {at={(1.05, 0.5)}, anchor = west, legend columns =1, draw=none, area legend},  ymajorgrids = true,width=0.83\textwidth,height=.35\textwidth, y label style={at={(axis description cs:-0.01,0.001)}},]%ytick=\empty]
\addplot[fill=blue,draw=black] table[x=Mx,y=My]{\shearyX};
\addlegendentry{MNIST}

\addplot[fill=red,draw=black] table[x=Fx,y=Fy]{\shearyX};
\addlegendentry{Fashion}

\addplot[fill=brown,draw=black] table[x=Ex,y=Ey]{\shearyX};
\addlegendentry{EMNIST}

\nextgroupplot[ybar, width=0.83\textwidth,height=.35\textwidth, bar width=.2cm,yshift=1cm,ymin=-10,ymax=0,legend style = {at={(1.05, 1.65)}, anchor = west, legend columns =1, draw=none, area    legend},ymajorgrids = true,xtick=data, xticklabels from table={\shearyY}{Algo}, xticklabel style={yshift=-2ex, xshift=-2ex,anchor= west},xlabel={Algorithms},x label style={at={(axis description cs:0.5,-0.1)},anchor=north}]
\addplot[fill=blue,draw=black] table[x=Mx,y=My]{\shearyY};
\addplot[fill=red,draw=black] table[x=Fx,y=Fy]{\shearyY};
\addplot[fill=brown,draw=black] table[x=Ex,y=Ey]{\shearyY};

\end{groupplot}
\end{tikzpicture}
\caption{Robustness graph for ShiftY property}
\end{figure}